Unsupervised embedding of trajectories captures the latent structure of mobility

### 2021-03-09 paper/current/main.tex
remove redundant para

%% word2vec
These trajectories are used as input to the standard skip-gram negative sampling word embedding, commonly known as \textit{word2vec}~\autocite{mikolov2013word2vec}.
\textit{word2vec} constructs dense and continuous vector representations of words and phrases, in which distance between words corresponds to a notion of semantic distance.
By embedding trajectories, we aim to learn a dense vector for every location, for which the distance between vectors relates to the tendency for two locations to occur in similar contexts.
Suppose a trajectory, denoted by ($a_{1}, a_{2}, \ldots, a_{T}$), where $a_{t}$ is the $t$th location in the trajectory. A location, $a_{t}$, is considered to have context locations, $a_{t-w}, \ldots, a_{t-1}, a_{t+1},\ldots, a_{t+w}$, that appear in the window surrounding $a_t$ up to a time lag of $w$, where $w$ is the window size parameter truncated at $t - w \geq 0$ and $t + w \leq T$. Then, the model learns probability $p(a_{t + \tau} \vert a_{t})$, where $-w\leq \tau\leq w$ and $\tau \neq 0$,  by maximizing its log likelihood given by
%
% MAIN W2V Equation
%
\begin{equation}
	\frac{1}{T}\sum_{t = 1}^{T} \sum_{-w \leq \tau \leq w, \tau \neq 0} \log p(a_{t + \tau} \vert a_{t}),
\end{equation}
where,

@@

There is an alternative implementation of \textit{word2vec} based on a matrix factorization~\autocite{levy2014neural,Qiu2018}, which optimizes the same objective function as for \texttt{gensim} but in a different way.
The matrix factorization approach generates embedding vectors such that the dot similarity for locations $i$ and $j$ is as close as $q_{ij}= \log(T_{ij} / s_i s_j) + \text{const}$.
An issue is that the generated embedding vectors can be largely affected by small flows $T_{ij}$ in the data.
For example, if there is no flow between locations $i$ and $j$ (\ie, $T_{ij} = 0$), then $q_{ij} = -\infty$, pushing $i$ and $j$ to the infinite horizon of the embedding space.
To remedy this problem, one can add a small amount of flow between such locations or exclude them from the embedding~\autocite{levy2014neural,Qiu2018}.
However, this may change the data to which we fit the model, and small flows may still have profound impact on the embedding space.
Therefore, we adopt the \texttt{gensim} implementation of the \textit{word2vec}.
Our simulations show that, in the prediction task of mobility flow, \texttt{gensim}'s word2vec performs better than the matrix factorization approach in xxx out of the xxx datasets (Fig.~??).

>>>>

%% word2vec
These trajectories are used as input to the standard skip-gram negative sampling word embedding, commonly known as \textit{word2vec}~\autocite{mikolov2013word2vec}.
\textit{word2vec} constructs dense and continuous vector representations of words and phrases, in which distance between words corresponds to a notion of semantic distance.
By embedding trajectories, we aim to learn a dense vector for every location, for which the distance between vectors relates to the tendency for two locations to occur in similar contexts.
Suppose a trajectory, denoted by ($a_{1}, a_{2}, \ldots, a_{T}$), where $a_{t}$ is the $t$th location in the trajectory. A location, $a_{t}$, is considered to have context locations, $a_{t-w}, \ldots, a_{t-1}, a_{t+1},\ldots, a_{t+w}$, that appear in the window surrounding $a_t$ up to a time lag of $w$, where $w$ is the window size parameter truncated at $t - w \geq 0$ and $t + w \leq T$. Then, the model learns probability $p(a_{t + \tau} \vert a_{t})$, where $-w\leq \tau\leq w$ and $\tau \neq 0$,  by maximizing its log likelihood given by
%
% MAIN W2V Equation
%
\begin{equation}
	{\cal J} = \frac{1}{T}\sum_{t = 1}^{T} \sum_{-w \leq \tau \leq w, \tau \neq 0} \log p(a_{t + \tau} \vert a_{t}),
\end{equation}
where,

@@
